---
title: "Homework 4"
author: "Adeline Shin"
date: "11/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gee)
library(geepack)
library(data.table)
library(lme4)
library(doBy)
library(tidyr)
library(mice)
library(purrr)
library(mitml)
library(CRTgeeDR)
```

# Uploading the Data
```{r}
# Upload the toenail df
toenail_df = read.delim("./toenail.txt", header = TRUE, sep = " ", col.names = c("Subject ID", "Response", "Treatment", "Month", "Visit", "X1", "X2", "X3"))

# Clean up toenail_df
toenail_df =
    toenail_df[, 1:5] %>% 
    janitor::clean_names()

toenail_df = as.data.table(toenail_df)
toenail_df$treatment = as.factor(toenail_df$treatment)
```

# Question 1
## Part 1
```{r}
# Add response at lag 1
toenail_df[, response_1 := shift(response, n = 1, type = "lag", fill = NA), by = "subject_id"]

# Transition Probabilities
tab1 = table(toenail_df$response, toenail_df$response_1)
tab1
round(prop.table(tab1, margin = 1), 2)

# Association b/w treatment and response
tab2 <- table(toenail_df$response, toenail_df$treatment)
round(prop.table(tab2, margin = 1), 2)

# Stratified by previous response
temp <- split(toenail_df, toenail_df$response_1)
tab3 <- lapply(temp, function(z){table(z$response, z$treatment)})
lapply(tab3, function(z){round(prop.table(z,margin = 1),2)})

# Model
model_lag_1 <- gee(response ~ treatment * response_1, corstr = "independence",family = binomial("logit"), id = subject_id, data = toenail_df)
round(summary(model_lag_1)$coeff,2)

model_lag_1b <- gee(response ~ treatment + response_1, corstr = "independence",family = binomial("logit"), id = subject_id, data = toenail_df)
round(summary(model_lag_1b)$coeff,2)
```

## Part 2
```{r}
toenail_df2 = toenail_df[, response_1 := shift(response, n = 2, type = "lag", fill = NA), by = "subject_id"]

# Transition Probabilities
tab1 = table(toenail_df$response, toenail_df$response_1)
tab1
round(prop.table(tab1, margin = 1), 2)

# Association b/w treatment and response
tab2 <- table(toenail_df$response, toenail_df$treatment)
round(prop.table(tab2, margin = 1), 2)

# Stratified by previous response
temp <- split(toenail_df, toenail_df$response_1)
tab3 <- lapply(temp, function(z){table(z$response, z$treatment)})
lapply(tab3, function(z){round(prop.table(z,margin = 1),2)})

# Model
model_lag_1 <- gee(response ~ treatment * response_1, corstr = "independence",family = binomial("logit"), id = subject_id, data = toenail_df)
round(summary(model_lag_1)$coeff,2)

model_lag_1b <- gee(response ~ treatment + response_1, corstr = "independence",family = binomial("logit"), id = subject_id, data = toenail_df)
round(summary(model_lag_1b)$coeff,2)
```

## Part 3
The following interpretations can be made for the coefficients in the first-order transition model: 

* Intercept: -3.41 is the log odds of having moderate or severe onycholysis for those who had no treatment and did not have moderate or severe onycholysis in the previous visit.
* treatment: -0.84 is the log odds ratio of respiratory infection comapring those with and without moderate or severe onycholysis who had an identical treatment status in the previous visit.
* response_1: 4.33 is the log odds ratio of moderate or severe onycholysis comparing those with and without treatment in their previous visit who currently have an identical status for moderate or severe onycholysis.

## Part 4
The interpretations from the transition model differ from the ones in HW 2 and HW 3 because they compare the current status to the status at the previous visit, while the previous models (GEE and mixed effects) do not account for the direct previous visit.

# Question 2
## Setup
```{r}
toenail <- fread("toenail.txt")
colnames(toenail) <- c("id","response","treatment","month","visit")
toenail2 <- tidyr::complete(toenail, id, visit) %>%
tidyr::fill(treatment)
toenail2 <- as.data.table(toenail2)
```

## Part 1
```{r warning = FALSE}
# Complete case analysis 
count = toenail2[, j = list(n = sum(!is.na(response))), by = "id"]
table(count$n)
count = count[n==7]
toenail3 = toenail2[id %in% count$id]
table(toenail3$response, useNA = "always")
table(toenail3$visit, toenail3$response, useNA = "always")
gee1 = geeglm(response ~ treatment + (visit + I(visit^2)) , id = id, data = toenail3, family = binomial(link = "logit"), corstr = "unstructured")
summary(gee1)
```

## Part 2
```{r warning = FALSE}
# Available case analysis 
table(toenail2$response,useNA = "always")
table(toenail2$visit, toenail2$response, useNA = "always")
gee2 = geeglm(response ~ treatment + (visit + I(visit^2)) , id = id, data = toenail2, family = binomial(link = "logit"), corstr = "unstructured")
summary(gee2)
```

## Part 3
```{r}
# LOCF
toenail4 = lapply(unique(toenail2$id), function(z){tidyr::fill(toenail2[id == z], treatment)})
toenail4 = rbindlist(toenail4)
table(toenail4$visit, toenail4$response, useNA = "always")
gee3 = geeglm(response ~ treatment + (visit + I(visit^2)), id = id, data = toenail4, family = binomial(link = "logit"), corstr = "unstructured")
summary(gee3)
```

## Part 4
```{r}
# MI 
toenail5 = toenail2
pred = make.predictorMatrix(toenail5)
pred 
pred["response", "id"] = -2
pred
pred = pred["response",,drop = FALSE]
pred
toenail5$id <- as.integer(toenail5$id)
imp = mice(toenail5, method = "2l.bin", pred = pred, seed = 1, maxit = 1, m = 5, print = FALSE, blocks = list(c("response")))
table(mice::complete(imp)$response, useNA = "always")

# GEE
implist = mids2mitml.list(imp)
gee4 = with(implist, geeglm(response ~ treatment + (visit + I(visit^2)), id = id, family = binomial, corstr = "unstructured"))
testEstimates(gee4)
```

## Part 5
```{r}
# ME
lme1 = mice::complete(imp, "all") %>%
    purrr::map(lme4::glmer,
               formula = response ~ treatment + (visit + I(visit^2)) + (1 | id),
               family = binomial) %>%
    pool()

summary(lme1)
```
